---
title: "R Notebook"
output: html_notebook
---

# ggplot2 is an open-source data visualization package for the statistical programming language R. 

```{r}
library(ggplot2)
```

```{r}
txhousing
```

Q1. Calculate the mean of the volume variable using the mean function without parameters. Assign it to Q1.
To access one variable of the dataset, use the $ operator, for example: txhousing$volume.

```{r}
Q1 <- mean(txhousing$volume)

Q1
```


Q2. Calculate the mean of the volume again. This time, use the parameter na.rm = TRUE.
Assign it to Q2


Compare Q1 and Q2. Ask yourself: What has changed? What is the difference?

R would not calculate the mean when there were missing values (NAs)

```{r}
Q2 <- mean(txhousing$volume, na.rm = TRUE)

Q2
```

Q3. Calculate the median of the listings variable. Assign it to Q3. Use what you have learned with the Q1 vs. Q2
comparison


```{r}
Q3 <- median(txhousing$listings, na.rm = TRUE)
  
Q3
```

Q4. Assign the min value of volume to Q4.


```{r}
Q4 <- min(txhousing$volume, na.rm = TRUE)

Q4
```

Q5. Assign the max value of volume to Q5.


```{r}
Q5 <- max(txhousing$volume, na.rm = TRUE)

Q5
```


Q6. Calculate the range of the volume. Assign it to Q6. (Hint: Use Q4 and Q5)


```{r}
Q6 <- Q5 - Q4

Q6
```

How many NAs are there in volume?

```{r}
is.na(txhousing$volume)
```


Use the sum function to count all the TRUE's from the is.na

```{r}
sum(is.na(txhousing$volume))
```

Find the proportion of NAs for the volume variable

There are packages in the DTSC 650 course that have functins to tell you this information. 
```{r}
sum(is.na(txhousing$volume)) / length(txhousing$volume)
```


Q7. Calculate the variance of the year variable. Assign it to Q7.

```{r}
Q7 <- var(txhousing$year)

Q7
```


Q8. Calculate the standard deviation of the volume variable. Assign it to Q8. 


```{r}
Q8 <- sd(txhousing$volume, na.rm = TRUE)

Q8
```

Q9. Using the summary function, create the summary of the txhousing dataset. Assign it to Q9.


```{r}
Q9 <- summary(txhousing)

Q9
```


Q10. Consider the first row of our dataset. Calculate the z-score of the volume of this observation (row). Assign it to Q10.
What does this z-score tell you about this observation?


```{r}
txhousing[1, ]
```
z score = (raw score - mean) / standard deviation

```{r}
Q10 <- (5380000 - mean(txhousing$volume, na.rm = TRUE)) / sd(txhousing$volume, na.rm = TRUE)

Q10

```

(5380000 - Q2) / Q8

A z-score of -0.4143106 means this house's volume was 0.4143106 standard deviations below the mean volume. 
